Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support).

### 2017-02-04 csf-2017/paper.tex
added part of intro

@@

\author{%
\IEEEauthorblockN{Lau Skorstengaard}
\IEEEauthorblockA{Aarhus University, Denmark\\
Email: lau@cs.au.dk} \and
\IEEEauthorblockN{Dominique~Devriese}
  \IEEEauthorblockA{iMinds-DistriNet, KU Leuven, Belgium\\
Email: dominique.devriese@cs.kuleuven.be} \and
\IEEEauthorblockN{Lars~Birkedal}\IEEEauthorblockA{Aarhus University, Denmark\\
Email: birkedal@cs.au.dk}}

@@

\section{Introduction}
\label{sec:introduction}

@@

\end{document}

>>>>

\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}

@@

\author{%
\IEEEauthorblockN{Lau~Skorstengaard}
\IEEEauthorblockA{Aarhus University, Denmark\\
Email: lau@cs.au.dk} \and
\IEEEauthorblockN{Dominique~Devriese}
  \IEEEauthorblockA{iMinds-DistriNet, KU Leuven, Belgium\\
Email: dominique.devriese@cs.kuleuven.be} \and
\IEEEauthorblockN{Lars~Birkedal}\IEEEauthorblockA{Aarhus University, Denmark\\
Email: birkedal@cs.au.dk}}

@@

\section{Introduction}
\label{sec:introduction}
%%% a)
% Current low-level protection coarse grained memory protection or
% properties of high-level languages are not really enforced?

% Capability machines offer fine-grained memory protection
A capability machine is a low-level machine that offers fine-grained
memory protection.
% Capability machines and unforgeable tokens for memory access. 
This is done by introducing unforgeable tokens to
the machine. These tokens are called capabillities and grant a kind of
authority and a range of authority. We will refer to the capabilities
that grant a combination of read, write, or execute permission as memory
capabilities. Capabilities with the necessary permissions must be
presented every time an instruction that manipulates the memory is
executed.

Memory capabilities provide memory protection, but they do not provide
a way to set up security domains. When you jump to an untrusted piece
of code that is supposed to jump back to you, then you have to either
revoke all your capabilities or pass to the piece of code you do not
trust. This short coming can be overcome by adding something like the
enter capability from the M-Machine\todo{add reference}. The enter
capability is an opaque capability which can only be used for a
jump. When jumped to it grants read and execute permission which
allows the code that is now executing to read capabilities stored in
memory. 
The CHERI processor's ccall achieves a similar property\todo{Add reference}.

Capabilities are irrevocable, so when we pass a capability has to
an untrusted program, then we have to assume that the program keeps it
around indefenitely. This means that we cannot reuse the piece of
memory that the capability governs essentially creating a memory
leak. The CHERI processor has a special kind of capabilities called
\emph{local capabilities}\todo{reference?} that introduces a simple
kind of temporal information control which allows for a simple kind of
capability revocation. This is achieved by adding a tag to every
capability that marks it as either local or global. Local capabilities
can only be written through a capability with a new permission called
\emph{permit write local}.
% More about local capabilities.

With just memory capabilities, enter capabailities and local
capabilities, the capability machine becomes very expressive and can
enforce properties that high-level languages promisse. The memory
capabilities enforce the privacy of local state as it is impossible
for anyone to access your local state without the capability to do
so. When this is combined with the enter capability,

%%% b)

% Examples of capability machines?

% Enter capabilities

% Local capabilities

% High-level languages often provide guarantees such as encapsulation of private state and well-bracketedness of calls - often not clear how well this is enforced. Just enforced when interacting with other programs written in the same high-level language or is it also guaranteed when interacting with assembly programs.

% High-level programs compiled to assembly can ensure these properties.

%

@@

\end{document}
\grid
